---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
<!-- README.md is generated from README.Rmd. Please edit that file -->

[![Build Status](https://travis-ci.org/phenomecentre/peakPantheR.svg?branch=master)](https://travis-ci.org/phenomecentre/peakPantheR)

# peakPantheR
Package for _Peak Picking and ANnoTation of High resolution Experiments in R_, implemented in `R` and `Shiny`


## Overview

`peakPantheR` implements functions to detect, integrate and report pre-defined features in MS files using XCMS3.
It is designed for:

* Real time feature detection and integration (_realTimeAnnotation_)
* Post-acquisition feature detection, integration and reporting (_parallelAnnotation_)


## Installation

Install the development version of the package directly from GitHub with:
```{r, eval = FALSE}
# Install devtools
if(!require("devtools")) install.packages("devtools")
devtools::install_github("phenomecentre/peakPantheR")
```

If the dependencies `xcms` and `MSnbase` are not successfully installed, `Bioconductor` must be added to the default repositories with:
```{r, eval = FALSE}
setRepositories(ind=1:2)
```


## Usage

**Real time compound integration concept:**

* process **multiple** compounds in **one** file at a time
* load list of expected _RT_ / _m/z_ regions of interest (ROI)
* detect features in each ROI and keep the highest intensity one
* determine peak statistics for each feature
* return a table with all detected compounds for that file (_row: compound, col: statistic_)

**Post-acquisition compound integration concept:**

* process **multiple** compounds in **multiple** files in **parallel**, store results in a **single** object
* load list of expected _RT_ / _m/z_ ROI and list of files to process
* initialise output object with expected ROI and file paths
* first pass (_without peak filling_):
    + for each file, detect features in each ROI and keep highest intensity
    + determine peak statistics for each feature
    + store results + EIC for each ROI
* visual inspection of first pass results, update ROI:
    + plot all EICs, peak apex _RT_ / _m/z_ & peak width evolution
    + correct ROI (remove interfering feature, correct _RT_ shift)
    + define fallback integration regions (FIR) if no feature is detected (median _RT_ / _m/z_ start and end of found features)
* initialise new output object, with updated regions of interest (uROI) and fallback integration regions (FIR)
* second pass (_with peak filling_):
    + for each file, detect features in each uROI and keep highest intensity
    + determine peak statistics for each feature
    + integrate FIR when no peaks found
    + store results + EIC for each uROI
* summary statistics:
    + plot EICs, apex and peakwidth evolution
    + compare first and second pass
* return result object and/or table (_row: file, col: compound_)


## Development, Code Structure, Documentation, Unit testing & Debugging

Required:

- [R](https://cran.rstudio.com/) (v3.4.3), install both 32bit and 64bit in `C:\R\R-3.4.3`
- [RTools](https://cran.r-project.org/bin/windows/Rtools/) (v3.4), install both 32bit and 64bit, even if only using x64. *Install in `C:\RTools`*

To install and modify `peakPantheR` (see also `dev_peakPantheR.R`):
```{r, eval = FALSE}
# Load devtools and roxygen2 for documentation
if(!require("devtools")) install.packages("devtools")
if(!require("roxygen2")) install.packages("roxygen2")

# Navigate to package dir
package_dir = 'path/peakPantheR'
setwd(package_dir)

# Generate documentation
devtools::document()

# Check the validity of the package (CRAN rules, should run unit tests)
devtools::check()

# Build package as .tar.gz for installation
devtools::build()

# Install and load the package on current machine
devtools::install()


# Remove package
remove.packages('peakPantheR')
#------------------------------------------------

# Load package
library(peakPantheR)
```

### Code Structure

Main functions are usually saved independently in `R\myFunction.R`, similar smaller functions can be grouped in a single file.

When calling functions from external packages, the package call must be explicit such as `pkg::fun()` for the function `fun()` in the package `pkg`.


### Documentation

Each function [documentation](http://r-pkgs.had.co.nz/man.html#man-functions) is generated using `roxygen`, which will parse [roxygen comments](http://r-pkgs.had.co.nz/man.html#roxygen-comments), add the function to the `Namespace` to make it available oustide the package (see `@export`), run the examples (see `@example`) and set links to other functions' documentation (see `@family`).
To regenerate the documentation, use `devtools::document()`


### Unit testing

Unit testing is achieved using the `testthat` package. By placing all tests in `inst/tests/` and one file in `test/` running them all, unit tests will be run during checks `devtools::check()`. 
In short, if the function is is `fun1.R` the corresponding tests are in `test_fun1.R`, and `run_tests.R` source the required functions and find all matching tests (see [a short introduction here](https://www.r-bloggers.com/unit-testing-with-r/amp/)). For more details, see [here](https://journal.r-project.org/archive/2011-1/RJournal_2011-1_Wickham.pdf), [here](http://r-pkgs.had.co.nz/tests.html) and [here (pg. 35-48)](http://www.is.uni-freiburg.de/ressourcen/algorithm-design-and-software-engineering-oeffentlicher-zugriff/11_softwaretesting.pdf).

> Tests need to run relatively quickly - aim for under a minute. Place skip_on_cran() at the beginning of long-running tests that shouldn't be run on CRAN - they'll still be run locally, but not on CRAN.


### Debugging

Interactive debugging is available in *RStudio*, more details can be found [here (pg.20-21)](http://www.is.uni-freiburg.de/ressourcen/algorithm-design-and-software-engineering-oeffentlicher-zugriff/11_softwaretesting.pdf).


## Copyright

`peakPantheR` is licensed under the [GPLv3](http://choosealicense.com/licenses/gpl-3.0/)

As a summary, the GPLv3 license requires attribution, inclusion of copyright and license information, disclosure of source code and changes. Derivative work must be available under the same terms.

&copy; National Phenome Centre (2018)


