---
title: "Getting Started with the peakPantheR package"
author: "Arnaud Wolfer"
date: "2018-02-21"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with the peakPantheR package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# peakPantheR
Package for _Peak Picking and ANnoTation of High resolution Experiments in R_, implemented in `R` and `Shiny`


## Overview

`peakPantheR` implements functions to detect, integrate and report pre-defined features in MS files using XCMS3.
It is designed for:

* **Real time** feature detection and integration (see [Real Time Annotation](real-time-annotation.html))
    + process `multiple` compounds in `one` file at a time
* **Post-acquisition** feature detection, integration and reporting (see [Parallel Annotation](parallel-annotation.html))
    + process `multiple` compounds in `multiple` files in `parallel`, store results in a `single` object


## Installation

Install the development version of the package directly from GitHub with:
```{r, eval = FALSE}
# Install devtools
if(!require("devtools")) install.packages("devtools")
devtools::install_github("phenomecentre/peakPantheR")
```

If the dependencies `xcms` and `MSnbase` are not successfully installed, `Bioconductor` must be added to the default repositories with:
```{r, eval = FALSE}
setRepositories(ind=1:2)
```


## Input Data

Both real time and parallel compound integration require a common set of information:

* Path(s) to `netCDF` / `mzML` MS file(s)
* An expected region of interest (`RT` / `m/z` window) for each compound.


The _`faahKO`_ package provides with a set of MS spectra that we can use for `peakPantheR` annotation:
```{r, echo = FALSE, include = FALSE}
# hide package load message
library(faahKO)
```
```{r}
library(faahKO)
## file paths
input_spectraPaths  <- c(system.file('cdf/KO/ko15.CDF', package = "faahKO"),
                         system.file('cdf/KO/ko16.CDF', package = "faahKO"),
                         system.file('cdf/KO/ko18.CDF', package = "faahKO"))
```
```{r, echo = FALSE}
library(faahKO)
input_spectraPaths  <- c(system.file('cdf/KO/ko15.CDF', package = "faahKO"),
                         system.file('cdf/KO/ko16.CDF', package = "faahKO"),
                         system.file('cdf/KO/ko18.CDF', package = "faahKO"))
input_spectraPaths
```

A table of targeted features contains as column:

* `cpdID` (numeric)
* `cpdName` (character)
* `rtMin` (sec)
* `rtMax` (sec)
* `rt` (sec, optional / `NA`)
* `mzMin` (m/z)
* `mzMax` (m/z)
* `mz` (m/z, optional / `NA`)

```{r, eval = FALSE}
# targetFeatTable
input_targetFeatTable     <- data.frame(matrix(vector(), 2, 8, dimnames=list(c(), c("cpdID", "cpdName", "rtMin", "rt", "rtMax", "mzMin", "mz", "mzMax"))), stringsAsFactors=F)
input_targetFeatTable[1,] <- c(1, "Cpd 1", 3310., 3344.888, 3390., 522.194778, 522.2, 522.205222)
input_targetFeatTable[2,] <- c(2, "Cpd 2", 3280., 3385.577, 3440., 496.195038, 496.2, 496.204962)
input_targetFeatTable[,c(1,3:8)] <- sapply(input_targetFeatTable[,c(1,3:8)], as.numeric)
```
```{r, results = "asis", echo = FALSE}
input_targetFeatTable     <- data.frame(matrix(vector(), 2, 8, dimnames=list(c(), c("cpdID", "cpdName", "rtMin", "rt", "rtMax", "mzMin", "mz", "mzMax"))), stringsAsFactors=F)
input_targetFeatTable[1,] <- c(1, "Cpd 1", 3310., 3344.888, 3390., 522.194778, 522.2, 522.205222)
input_targetFeatTable[2,] <- c(2, "Cpd 2", 3280., 3385.577, 3440., 496.195038, 496.2, 496.204962)
input_targetFeatTable[,c(1,3:8)] <- sapply(input_targetFeatTable[,c(1,3:8)], as.numeric)
rownames(input_targetFeatTable) <- NULL
pander::pandoc.table(input_targetFeatTable, digits = 9)
```


## See Also

* [Real Time Annotation](real-time-annotation.html)
* [Parallel Annotation](parallel-annotation.html)
